---
title: "Job Market exploration"
author: "Mahd Sardar"
date: "2023-04-15"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    theme: flatly
    highlight: monochrome
    code_folding: hide
    markdown_extensions:
      - admonition
    warnings: false
---

# Setup

```{r setup, include = FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	messages = FALSE,
	comment = "##",
	collapse = TRUE,
	fig.width = 40, 
	fig.height = 20,
	fig.path = "C:/Users/smf19/OneDrive/Desktop/Jobster/R/graphs",
	dev = "png"
)

# kable
options(knitr.kable.NA = '')

# Set the graphical theme
ggplot2::theme_set(ggplot2::theme_light())
library(ggplot2)
library(tidyverse)
library(stringr)
library(iai)
library(AER)
library(cowplot)
library(kableExtra)
library(gridExtra)
library(broom)
library(haven)
library(fixest)
library(stargazer)
# load model summary and set options
library(modelsummary)
gm <- modelsummary::gof_map
gm$omit <- TRUE
gm$omit[1] <- FALSE
gm$omit[6] <- FALSE
gm$omit[5] <- FALSE
gm$omit[17] <- FALSE

```


# Load Data

```{r load data}
my_theme <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size = 30,margin = margin(t = 10)),
        axis.text.y = element_text(vjust = 0.5, hjust=1, size = 30,margin = margin(r = 10)),
        axis.title.y = element_text(size = 30, margin = margin(r = 30)),
        axis.title.x = element_text(size = 30, margin = margin(t = 20)),
        panel.grid.major = element_line(color = "gray80"),
        panel.grid.minor = element_line(color = "gray90"),
        panel.background = element_rect(fill = "white"),
        axis.line = element_line(colour = "black"),
        legend.key.size = unit(10, "points"), 
        legend.text = element_text(size = 20, margin = margin(b = 10)),
        legend.title = element_text(size = 30),
        legend.position = "bottom",
        plot.title = element_text(size=60, margin = margin(t = 50 , b = 50)))

df <- read.csv("../rawdata/14100023.csv")

categorical_vars <- c("GEO", "Labour.force.characteristics","North.American.Industry.Classification.System..NAICS.", "Sex", "Age.group")

df[categorical_vars] <- lapply(df[categorical_vars], as.factor)


data_diff <- df %>%
  mutate(North.American.Industry.Classification.System..NAICS. = str_extract(North.American.Industry.Classification.System..NAICS., "^[^\\[]+")) %>%
  filter(REF_DATE >= 2000 & Labour.force.characteristics == "Unemployment rate") %>%
  select(-c(UOM, UOM_ID, SCALAR_FACTOR, SCALAR_ID, VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS,Labour.force.characteristics)) %>%
  drop_na()


df_productivity <- read.csv("..\\rawdata\\productivity data\\36100208.csv") %>%
  filter(Multifactor.productivity.and.related.variables %in% c("Labour productivity") & REF_DATE >= 2000 ) %>%
  mutate(across(c(GEO, North.American.Industry.Classification.System..NAICS.), as.factor))%>%
  rename(`Lp_index` = VALUE,# 2012 = 100
         Industry = North.American.Industry.Classification.System..NAICS.,
         Year = REF_DATE) %>% 
  select(-c(UOM, UOM_ID, SCALAR_FACTOR, SCALAR_ID, VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS,Multifactor.productivity.and.related.variables))


df_unionization <- read.csv("..\\rawdata\\unionization data\\14100132.csv") %>%
  filter(Union.status == "Unionization rate" & REF_DATE >= 2000) %>%
  mutate(across(c(GEO, Union.status, North.American.Industry.Classification.System..NAICS., Sex), as.factor)) %>%
  select(-c(UOM, UOM_ID, SCALAR_FACTOR, SCALAR_ID, VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS, Union.status))

df_tech <- read.csv("..\\rawdata\\ai data\\27100367.csv") %>%
  filter(REF_DATE >= 2000 &
           !North.American.Industry.Classification.System..NAICS. %in% c("Total, all industries", "Goods-producing sector", "Services-producing sector")) %>%
  mutate(across(c(GEO, Advanced.or.emerging.technologies), as.factor)) %>%
  mutate(North.American.Industry.Classification.System..NAICS. = str_extract(North.American.Industry.Classification.System..NAICS., "^[^\\[]+")) %>%
  select(-c(UOM, UOM_ID, SCALAR_FACTOR, SCALAR_ID, VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS))

df_gdp <- read.csv("..\\rawdata\\gdp data\\36100401.csv") %>%
  filter(REF_DATE >= 2000 & GEO == "Canada") %>%
  select(-c(VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS, UOM_ID, SCALAR_ID))
 

df_investment <- read.csv("..\\rawdata\\investment data\\34100035.csv") %>%
  select(-c(VECTOR, COORDINATE, DGUID, STATUS, SYMBOL, TERMINATED, DECIMALS, UOM_ID, SCALAR_ID)) %>%
  filter(Capital.and.repair.expenditures == "Capital expenditures" & GEO == "Canada") %>%
  rename(Industry = North.American.Industry.Classification.System..NAICS.,
         Year = REF_DATE
         )
```

# Visualize Unemployment trends

```{r visualize unemp trends}
unemployment_rate_data <- df %>%
  filter(Labour.force.characteristics == "Unemployment rate" & GEO == "Canada" & REF_DATE >= "2000" & Age.group == "25 to 54 years" & 
           Sex == "Both sexes") %>%
  group_by(REF_DATE, North.American.Industry.Classification.System..NAICS.) %>%
  drop_na(VALUE)

ggplot(unemployment_rate_data, aes(x = REF_DATE, y = VALUE, color = North.American.Industry.Classification.System..NAICS., group = North.American.Industry.Classification.System..NAICS.)) +
  geom_line(size = 2) + 
  scale_color_discrete(name = "") + # Change the name of the legend title
  labs( x= "Year", y = "Unemployment Rate", title = 'Unemployment Rate by Industry') +
  my_theme +
    geom_text(data = unemployment_rate_data %>% group_by(North.American.Industry.Classification.System..NAICS.) %>% summarize(REF_DATE = max(REF_DATE), VALUE = max(VALUE)), aes(x = REF_DATE, y = VALUE, label = North.American.Industry.Classification.System..NAICS.),hjust = 0.1, size = 6, check_overlap = TRUE)
```


# Visualize Sectorized Trends (Only Goods and Service Sectors)

```{r visualize sectorized trends}
GoodsSector <- factor(c("Agriculture", "Forestry, fishing, mining, quarrying, oil and gas", "Utilities", "Construction", "Manufacturing"))

ServicesSector <- factor(c("Wholesale and retail trade", "Transportation and warehousing", "Finance, insurance, real estate, rental and leasing",
                     "Professional, scientific and technical services", "Business, building and other support services", "Educational services",
                     "Health care and social assistance", "Information, culture and recreation", "Accommodation and food services",
                     "Other services \\(except public administration\\)", "Public administration"))

all_sectors <- c(GoodsSector, ServicesSector)

unemp_sectorized <- unemployment_rate_data %>%
  mutate(Industry = factor(str_extract(North.American.Industry.Classification.System..NAICS., "^[^\\[]+"))) %>%
  mutate(Sector = ifelse(str_detect(Industry, paste(GoodsSector, collapse = "|")), "Goods", "Services")) %>%
  filter(str_detect(Industry, paste(all_sectors, collapse = "|")))


ggplot(unemp_sectorized, aes(x = REF_DATE, y = VALUE, color = Industry, group = Industry)) +
  geom_line(size = 2) + 
  labs(x = "Year", y = "Unemployment Rate", title = 'Unemployment Rate by Industry (Goods/Service Sectors Only)') +
  scale_color_discrete(name = "") + # Change the name of the legend title
  my_theme +
    geom_text(data = unemp_sectorized %>% 
                group_by(Industry) %>% 
                summarize(REF_DATE = max(REF_DATE), VALUE = max(VALUE)), 
              aes(x = REF_DATE, y = VALUE, label = Industry),hjust = 0.1, size = 6, check_overlap = TRUE)
```

# Unemployment trends per industry

```{r unemployment trend per industry}
ggplot(unemp_sectorized, aes(x = REF_DATE, y = VALUE, color = Industry, group = Industry)) +
  geom_line(size = 2) + 
  labs(x = "Year", y = "Unemployment Rate", title = 'Unemployment Rate per Industry') +
  facet_wrap(~ Industry, scales = "free_y", ncol = 4) +
  my_theme
```

# Correlation matrix to assess Differences-in-Differences assumption

```{r assess parallel trends (DiD)}
cor_data <- unemp_sectorized %>%
  select(REF_DATE, North.American.Industry.Classification.System..NAICS., VALUE) %>%
  spread(North.American.Industry.Classification.System..NAICS., VALUE) %>%
  select(-REF_DATE) %>%
  cor()

cor_data[lower.tri(cor_data)] <- NA
```


# Refactor Data (for futher analysis)

```{r reform data}
df_lm <- df %>%
  filter(Labour.force.characteristics == "Unemployment rate" & GEO == "Canada"  & REF_DATE >= 2000) %>%
  select(-c("Labour.force.characteristics")) 

df_lm$Age.group <- droplevels(df_lm$Age.group, exclude = "15 years and over")


merged_union <- full_join(df_lm, df_unionization, by = c("REF_DATE", "North.American.Industry.Classification.System..NAICS.", "GEO", "Sex")) %>%
  select(-c(SYMBOL, TERMINATED, STATUS, UOM, UOM_ID, SCALAR_ID, SCALAR_FACTOR, VECTOR, COORDINATE, DECIMALS, DGUID)) %>%
  drop_na()

merged_union$Sex <- droplevels(merged_union$Sex, exclude = "Both sexes")


merge_UandGDP <- full_join(merged_union, df_gdp, by = c("REF_DATE", "North.American.Industry.Classification.System..NAICS.")) %>%
  rename(Unemp = VALUE.x,
         Union = VALUE.y,
         GDP= VALUE,
         Industry = North.American.Industry.Classification.System..NAICS.,
         Year = REF_DATE
         ) %>%
  select(-c(GEO.y, SCALAR_FACTOR, UOM)) %>%
  drop_na()

merge_UIGDP <- full_join(merge_UandGDP, df_investment, by = c("Year" , "Industry")) %>%
  select(-c(`SCALAR_FACTOR`, Capital.and.repair.expenditures, GEO.x, UOM)) %>%
  rename(Capex = VALUE) %>%
  drop_na()


merge_ULPGDP <- full_join(merge_UIGDP, df_productivity, by = c("Year","Industry")) %>%
  select(-c(GEO.y)) %>%
  drop_na()

merge_OnlyLp <- full_join(merge_UandGDP, df_productivity, by = c("Year", "Industry")) %>%
  select(-c(GEO.x)) %>%
  drop_na()
```


# Check Skewness

```{r check skewness}
# distributions
par(mfrow = c(3,2))

# Pre-transformation
hist(merge_UIGDP$Unemp, main = "Unemployment")
hist(merge_UIGDP$GDP, main = "GDP")
hist(merge_UIGDP$GDP, main = "GDP")
hist(merge_UIGDP$Capex, main = "Capital Expenditure")
hist(merge_ULPGDP$Lp_index, main = "Labour Productivity Index")


# log-transformed
hist(log(merge_UIGDP$Unemp),  main = "Log Unemployment")
hist(log(merge_UIGDP$GDP), main = "Log GDP")
hist(log(merge_UIGDP$Capex), main = "Log Capital Expenditure")
```


# Build Stratified Model

```{r stratified model}
# Create a list of industry, age group, and sex classification combinations
industries <- unique(merge_UIGDP$Industry)
ages <- factor(unique(merge_UIGDP$Age.group), exclude = "15 years and over")
sexes <- factor(unique(merge_UIGDP$Sex), exclude = "Both sexes")

combo_list <- expand.grid(Industry = industries, Age_Group = ages, Sex = sexes) %>%
  drop_na()



reg_func <- function(df) {
  lm(log(Unemp) ~ Union + log(GDP) + log(Capex) , data = df)
}

# Use lapply to apply the function to each combination of variables
reg_results <- lapply(1:nrow(combo_list), function(i) {
  industry <- combo_list$Industry[i]
  age <- combo_list$Age_Group[i]
  sex <- combo_list$Sex[i]
  df <- merge_UIGDP %>%
    filter(Industry == industry & Age.group == age & Sex == sex)
  if (nrow(df) > 0) {
    reg_result <- reg_func(df)
    reg_result$industry <- industry
    reg_result$age_group <- age
    reg_result$sex <- sex
    reg_result
  } else {
    NULL
  }
})


# Check for null models
null_indices <- which(sapply(reg_results, is.null))
# data frame that holds all NULL combinations
combo_list[null_indices, ]
# remove NULL models
reg_results <- reg_results[!sapply(reg_results, is.null)]


# Combine the results into a single dataframe
reg_df <- as.data.frame(do.call(rbind, reg_results))

# List of modelsummary tables
modsummaries <- lapply(reg_results, function(model) {
  if (!is.null(model)) {
    modelsummary(model, 
                 statistic_override = sandwich,
                 stars = T,
                 fmt = 8,
                 output = "html")
  } else {
    NULL
  }
})

```

# Generate Stratified Table

```{r generate stratified table}
# Use nest() function to group the data by Industry, Age_Group, and Sex
grouped_data <- merge_UIGDP %>% 
  group_by(Industry, Age.group, Sex) %>% 
  nest()

# Use map() function to apply the reg_func() to each group
reg_results_grouped <- grouped_data %>%
  mutate(model = map(data, reg_func)) %>%
  filter(!map_lgl(model, ~ any(is.na(coef(.x))) | any(is.nan(coef(.x))))) %>%
  mutate(n_obs = sapply(data, nrow))


# Extract the coefficients, industry, age group, and sex for each model

# Define a function to format p-values
format_p_value <- function(p) {
  if (is.na(p)) {
    return("")
  } else if (p < 0.001) {
    return(cell_spec(paste0(format(p, nsmall = 6), "***"), color = "green", bold = TRUE, underline = TRUE))
  } else if (p < 0.01) {
    return(cell_spec(paste0(format(p, nsmall = 6), "**"), color = "green", bold = TRUE))
  } else if (p < 0.05) {
    return(cell_spec(paste0(format(p, nsmall = 6), "*"), color = "green"))
  } else {
    return(format(p, nsmall = 6))
  }
}


results_table_filtered <- reg_results_grouped %>%
  mutate(intercept = map_dbl(model, ~coef(.)[1]),
         unionization_rate = map_dbl(model, ~coef(.)[2]),
         gdp = map_dbl(model, ~coef(.)[3]),
         capex = map_dbl(model, ~coef(.)[4]),
         p_union = map_dbl(model, ~coeftest(., vcov = vcovHC(., type = 'HC1'))[2, 4]),
         p_gdp = map_dbl(model, ~coeftest(., vcov = vcovHC(., type = 'HC1'))[3, 4]),
         p_capex = map_dbl(model, ~coeftest(., vcov = vcovHC(., type = 'HC1'))[4, 4]),
         durbin_watson = map_dbl(model, ~durbinWatsonTest(.)$dw),
         p_dw = map_dbl(model, ~durbinWatsonTest(.)$p)) %>%
  select(intercept, unionization_rate, gdp, capex, p_union, p_gdp, p_capex, durbin_watson, p_dw) %>%
  mutate(across(everything(), round, 6),
         p_union = format_p_value(p_union),
         p_gdp = format_p_value(p_gdp),
         p_capex = format_p_value(p_capex),
         p_dw = format_p_value(p_dw)) %>%
  filter(p_union < 0.05 | p_gdp < 0.05 | p_capex < 0.05) %>%
  mutate(across(c(p_union, p_gdp, p_capex, p_dw), ~ if_else(. == "", "", paste0(., " ")))) %>%
  mutate(across(c(p_union, p_gdp, p_capex, p_dw), as_factor)) %>%
  kable(format = "html", 
        caption = "Regression Results",
        escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), 
                full_width = FALSE, 
                fixed_thead = TRUE,
                position = "center") %>%
  column_spec(5:7, bold = TRUE)

results_table_filtered
```

# Fixed Effects Model Specifications

```{r fixed effects models}
# Fixed effects model
FE1 <- feols(Unemp ~ log(Capex) | Industry + Age.group + Sex, merge_UIGDP, cluster = merge_UIGDP$Industry)
FE2 <- feols(Unemp ~ Union| Industry + Age.group + Sex, merge_UIGDP, cluster = merge_UIGDP$Industry)
FE3 <- feols(Unemp ~ log(GDP)| Industry + Age.group + Sex, merge_UIGDP, cluster = merge_UIGDP$Industry)
FE4 <- feols(Unemp ~ Union + log(GDP) + log(Capex) | Industry + Age.group + Sex, merge_UIGDP,cluster = merge_UIGDP$Industry)
FE5 <- feols(Unemp ~ log(GDP) + Union + Lp_index | Industry + Age.group + Sex, merge_ULPGDP, cluster = merge_ULPGDP$Industry)


etable(FE1,FE2,FE3, FE4,FE5,
      headers = c("log(Capex)", "Unionization rate","log(GDP)", "Full", "Full+Lp"))
```


# Fixed Effects Heteroscedasticity Test

```{r FE model heteroscedasticity}
# Test FE model errors for heteroscedasticity

# Create a data frame with the fitted values and residuals
FE1_data <- data.frame(fitted = FE1$fitted.values, resid = FE1$residuals)
FE2_data <- data.frame(fitted = FE2$fitted.values, resid = FE2$residuals)
FE3_data <- data.frame(fitted = FE3$fitted.values, resid = FE3$residuals)
FE4_data <- data.frame(fitted = FE4$fitted.values, resid = FE4$residuals)

# Plot the fitted values versus residuals for each model

grid.arrange(
  ggplot(FE1_data, aes(x = fitted, y = resid)) +
    geom_point() +
    geom_smooth(se = TRUE) + 
    my_theme +
    ggtitle("FE1: Fitted vs Residuals"),
    
  ggplot(FE2_data, aes(x = fitted, y = resid)) +
    geom_point() +
    geom_smooth(se = TRUE) +
    my_theme +
    ggtitle("FE2: Fitted vs Residuals"),
    
  ggplot(FE3_data, aes(x = fitted, y = resid)) +
    geom_point() +
    geom_smooth(se = TRUE) +
    my_theme +
    ggtitle("FE3: Fitted vs Residuals"),
    
  ggplot(FE4_data, aes(x = fitted, y = resid)) +
    geom_point() +
    geom_smooth(se = TRUE) +
    my_theme +
    ggtitle("FE4: Fitted vs Residuals"),
  
  ncol = 2
)
#bp tests
bptest(FE1$residuals ~ FE1$fitted.values^2)
bptest(FE2$residuals ~ FE2$fitted.values^2)
bptest(FE3$residuals ~ FE3$fitted.values^2)
bptest(FE4$residuals ~ FE3$fitted.values^2)
```


# Linear Relationship Assumption

```{r linear relationship assumption}
# Linear relationship assumption
# Set up custom color scheme
my_colors <- c("#E41A1C", "#377EB8", "#4DAF4A")

grid.arrange(
  # Scatterplot of Unemp vs. log(GDP)

  ggplot(merge_UIGDP, aes(x = log(GDP), y = Unemp)) +
    geom_point(color = my_colors[1]) +
    geom_smooth(method = "lm", se = TRUE, color = my_colors[1]) +
    labs(x = "Log of Gross Domestic Product (GDP)", y = "Log of Unemployment Rate") +
    ggtitle("Scatterplot of Unemp vs. Log(GDP)") +
    my_theme,
  
  # Scatterplot of Unemp vs. log(Capex)
  ggplot(merge_UIGDP, aes(x = log(Capex), y = Unemp)) +
    geom_point(color = my_colors[2]) +
    geom_smooth(method = "lm", se = TRUE, color = my_colors[2]) +
    labs(x = "Log of Capital Expenditure (Capex)", y = "Log of Unemployment Rate") +
    ggtitle("Scatterplot of Unemp vs. Log(Capex)") +
    my_theme,
  
  # Scatterplot of Unemp vs. Union
  ggplot(merge_UIGDP, aes(x = Union, y = Unemp)) +
    geom_point(color = my_colors[3]) +
    geom_smooth(method = "lm", se = TRUE, color = my_colors[3]) +
    labs(x = "Union Membership (in Percent)", y = "Log of Unemployment Rate") +
    ggtitle("Scatterplot of Unemp vs. Union Membership") +
    my_theme,
  
  ncol = 2

)
```

# Helper functions

```{r helper functions}
# Helper functions

# function to find model index given specific combination of control variables

find_model <- function(NAICS, AGES, SEX) {
  # Loop through reg_results and check if combination matches input values
  for (i in 1:length(reg_results)) {
    if (reg_results[[i]]$industry == NAICS &
        reg_results[[i]]$age_group == AGES &
        reg_results[[i]]$sex == SEX) {
      return(i)
    }
  }
  # If no matching model is found, return NA
  return(NA)
}


# Set values for NAICS, AGES, and SEX

NAICS_val <- "Utilities [22]"
AGES_val <- "15 years and over"
SEX_val <- "Both sexes"

# Find index of corresponding model

model_index <- find_model(NAICS_val, AGES_val, SEX_val)

# Print lm object for corresponding model

if (!is.na(model_index)) {
  print(reg_results[[model_index]])
} else {
  print("No matching model found.")
}

# Inverse function that returns combination given model index

get_model_vars <- function(index, data = reg_results) {
  industry <- reg_results[[index]]$industry
  age_group <- reg_results[[index]]$age_group
  sex <- reg_results[[index]]$sex
  return(data.frame(Industry = industry, Age_Group = age_group, Sex = sex))
} 

get_model_vars(1)
```







